---
title: "Lab_7"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install(version = "3.11")
```

```{r}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("rnaseqGene")

# I also needed to install the following on my computer 
# (You will realize this if you get an error message that the corresponding library will not load)

BiocManager::install("airway")
BiocManager::install("tximeta")
BiocManager::install("DESeq2")
BiocManager::install("Gviz")
BiocManager::install("sva")
BiocManager::install("RUVSeq")
BiocManager::install("fission")
```

```{r}
library("airway")
```

```{r}
dir <- system.file("extdata", package="airway", mustWork=TRUE)
```

```{r}
list.files(dir)
```

```{r}
list.files(file.path(dir, "quants"))
```

```{r}
csvfile <- file.path(dir, "sample_table.csv")
coldata <- read.csv(csvfile, row.names=1, stringsAsFactors=FALSE)
coldata
```

```{r}
coldata <- coldata[1:2,]
coldata$names <- coldata$Run
coldata$files <- file.path(dir, "quants", coldata$names, "quant.sf.gz")
file.exists(coldata$files)
```

```{r}
library("tximeta")
se <- tximeta(coldata)

```
```{r}
dim(se)
```

```{r}
head(rownames(se))

```

```{r}
gse <- summarizeToGene(se)
```

```{r}
dim(gse)
```

```{r}
head(rownames(gse))
```

```{r}
data(gse)
gse
```

```{r}
assayNames(gse)
```

```{r}
head(assay(gse), 3)
```

```{r}
colSums(assay(gse))
```

```{r}
rowRanges(gse)
```

```{r}
seqinfo(rowRanges(gse))
```

```{r}
colData(gse)
```

```{r}
gse$donor
```

```{r}
gse$condition
```

```{r}
gse$cell <- gse$donor
gse$dex <- gse$condition
```

```{r}
levels(gse$dex)
```

```{r}
# when renaming levels, the order must be preserved!
levels(gse$dex) <- c("untrt", "trt")

```

### it is prefered in R that the first level of a factor be the reference level (e.g. control, or untreated samples).

```{r}
library("magrittr")
gse$dex %<>% relevel("untrt")
gse$dex
```

### %<>% is the compound assignment pipe-operator from the magrittr package, the above line of code is a concise way of saying:

```{r}
gse$dex <- relevel(gse$dex, "untrt")
```

### the second argument of round tells how many decimal points to keep

```{r}
round( colSums(assay(gse)) / 1e6, 1 )
```
```{r}
library("DESeq2")
```

```{r}
dds <- DESeqDataSet(gse, design = ~ cell + dex)
```

```{r}
countdata <- round(assays(gse)[["counts"]])
head(countdata, 3)
```

### If you’ve imported the count data in some other way, for example loading a pre-computed count matrix, it is very important to check manually that the columns of the count matrix correspond to the rows of the sample information table.
 
```{r}
coldata <- colData(gse)
```

```{r}
ddsMat <- DESeqDataSetFromMatrix(countData = countdata,
                                 colData = coldata,
                                 design = ~ cell + dex)
```

```{r}
nrow(dds)
```

```{r}
keep <- rowSums(counts(dds)) > 1
dds <- dds[keep,]
nrow(dds)
```

```{r}
# at least 3 samples with a count of 10 or higher
keep <- rowSums(counts(dds) >= 10) >= 3
```

```{r}
lambda <- 10^seq(from = -1, to = 2, length = 1000)
cts <- matrix(rpois(1000*100, lambda), ncol = 100)
library("vsn")
meanSdPlot(cts, ranks = FALSE)
```

```{r}
log.cts.one <- log2(cts + 1)
meanSdPlot(log.cts.one, ranks = FALSE)
```
```{r}
vsd <- vst(dds, blind = FALSE)
head(assay(vsd), 3)
```

```{r}
colData(vsd)
```

```{r}
rld <- rlog(dds, blind = FALSE)
head(assay(rld), 3)

```

### In the above function calls, we specified blind = FALSE, which means that differences between cell lines and treatment (the variables in the design) will not contribute to the expected variance-mean trend of the experiment.

```{r}
library("dplyr")
library("ggplot2")

dds <- estimateSizeFactors(dds)

df <- bind_rows(
  as_data_frame(log2(counts(dds, normalized=TRUE)[, 1:2]+1)) %>%
         mutate(transformation = "log2(x + 1)"),
  as_data_frame(assay(vsd)[, 1:2]) %>% mutate(transformation = "vst"),
  as_data_frame(assay(rld)[, 1:2]) %>% mutate(transformation = "rlog"))
  
colnames(df)[1:2] <- c("x", "y")  

lvls <- c("log2(x + 1)", "vst", "rlog")
df$transformation <- factor(df$transformation, levels=lvls)

ggplot(df, aes(x = x, y = y)) + geom_hex(bins = 80) +
  coord_fixed() + facet_grid( . ~ transformation)  
```

```{r}
sampleDists <- dist(t(assay(vsd)))
sampleDists
```

```{r}
library("pheatmap")
library("RColorBrewer")
```

```{r}
sampleDistMatrix <- as.matrix( sampleDists )
rownames(sampleDistMatrix) <- paste( vsd$dex, vsd$cell, sep = " - " )
colnames(sampleDistMatrix) <- NULL
colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
pheatmap(sampleDistMatrix,
         clustering_distance_rows = sampleDists,
         clustering_distance_cols = sampleDists,
         col = colors)
```

### Note that we have changed the row names of the distance matrix to contain treatment type and patient number instead of sample ID, so that we have all this information in view when looking at the heatmap.

```{r}
library("PoiClaClu")
poisd <- PoissonDistance(t(counts(dds)))
```

```{r}
samplePoisDistMatrix <- as.matrix( poisd$dd )
rownames(samplePoisDistMatrix) <- paste( dds$dex, dds$cell, sep=" - " )
colnames(samplePoisDistMatrix) <- NULL
pheatmap(samplePoisDistMatrix,
         clustering_distance_rows = poisd$dd,
         clustering_distance_cols = poisd$dd,
         col = colors)
```

### Another way to visualize sample-to-sample distances is a principal components analysis (PCA). In this ordination method, the data points (here, the samples) are projected onto the 2D plane such that they spread out in the two directions that explain most of the differences

```{r}
plotPCA(vsd, intgroup = c("dex", "cell"))
```

```{r}
pcaData <- plotPCA(vsd, intgroup = c( "dex", "cell"), returnData = TRUE)
pcaData
```

```{r}
percentVar <- round(100 * attr(pcaData, "percentVar"))
```

```{r}
ggplot(pcaData, aes(x = PC1, y = PC2, color = dex, shape = cell)) +
  geom_point(size =3) +
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
  coord_fixed() +
  ggtitle("PCA with VST data")
```
### Here we specify cell line (plotting symbol) and dexamethasone treatment (color).

```{r}
library("glmpca")
gpca <- glmpca(counts(dds), L=2)
gpca.dat <- gpca$factors
gpca.dat$dex <- dds$dex
gpca.dat$cell <- dds$cell
```

```{r}
ggplot(gpca.dat, aes(x = dim1, y = dim2, color = dex, shape = cell)) +
  geom_point(size =3) + coord_fixed() + ggtitle("glmpca - Generalized PCA")
```

```{r}
mds <- as.data.frame(colData(vsd))  %>%
         cbind(cmdscale(sampleDistMatrix))
ggplot(mds, aes(x = `1`, y = `2`, color = dex, shape = cell)) +
  geom_point(size = 3) + coord_fixed() + ggtitle("MDS with VST data")
```

```{r}
mdsPois <- as.data.frame(colData(dds)) %>%
   cbind(cmdscale(samplePoisDistMatrix))
ggplot(mdsPois, aes(x = `1`, y = `2`, color = dex, shape = cell)) +
  geom_point(size = 3) + coord_fixed() + ggtitle("MDS with PoissonDistances")

```
```{r}
dds <- DESeq(dds)
```

```{r}
res <- results(dds)
res
```

```{r}
res <- results(dds, contrast=c("dex","trt","untrt"))
```

```{r}
mcols(res, use.names = TRUE)
```

```{r}
summary(res)
```

### If we lower the false discovery rate threshold, we should also inform the results() function about it, so that the function can use this threshold for the optimal independent filtering that it performs:

```{r}
res.05 <- results(dds, alpha = 0.05)
table(res.05$padj < 0.05)
```

```{r}
resLFC1 <- results(dds, lfcThreshold=1)
table(resLFC1$padj < 0.1)
```

### In general, the results for a comparison of any two levels of a variable can be extracted using the contrast argument to results.

```{r}
results(dds, contrast = c("cell", "N061011", "N61311"))
```

```{r}
sum(res$pvalue < 0.05, na.rm=TRUE)
```

```{r}
sum(!is.na(res$pvalue))
```

```{r}
sum(res$pvalue < 0.05, na.rm=TRUE)
```

```{r}
sum(!is.na(res$pvalue))
```

```{r}
sum(res$padj < 0.1, na.rm=TRUE)

```

### We subset the results table to these genes and then sort it by the log2 fold change estimate to get the significant genes with the strongest down-regulation:

```{r}
resSig <- subset(res, padj < 0.1)
head(resSig[ order(resSig$log2FoldChange), ])
```

### strongest up regulation

```{r}
head(resSig[ order(resSig$log2FoldChange, decreasing = TRUE), ])
```


### A quick way to visualize the counts for a particular gene is to use the plotCounts function that takes as arguments the DESeqDataSet, a gene name, and the group over which to plot the counts (figure below).

```{r}
topGene <- rownames(res)[which.min(res$padj)]
plotCounts(dds, gene = topGene, intgroup=c("dex"))

```

```{r}
library("ggbeeswarm")
geneCounts <- plotCounts(dds, gene = topGene, intgroup = c("dex","cell"),
                         returnData = TRUE)
ggplot(geneCounts, aes(x = dex, y = count, color = cell)) +
  scale_y_log10() +  geom_beeswarm(cex = 3)
```
```{r}
ggplot(geneCounts, aes(x = dex, y = count, color = cell, group = cell)) +
  scale_y_log10() + geom_point(size = 3) + geom_line()
```
### Note that the DESeq test actually takes into account the cell line effect, so this figure more closely depicts the difference being tested.

```{r}
library("apeglm")
resultsNames(dds)
```

```{r}
res <- lfcShrink(dds, coef="dex_trt_vs_untrt", type="apeglm")
plotMA(res, ylim = c(-5, 5))
```

### An MA-plot (Dudoit et al. 2002) provides a useful overview for the distribution of the estimated coefficients in the model, e.g. the comparisons of interest, across all genes

```{r}
res.noshr <- results(dds, name="dex_trt_vs_untrt")
plotMA(res.noshr, ylim = c(-5, 5))
```

```{r}
plotMA(res, ylim = c(-5,5))
topGene <- rownames(res)[which.min(res$padj)]
with(res[topGene, ], {
  points(baseMean, log2FoldChange, col="dodgerblue", cex=2, lwd=2)
  text(baseMean, log2FoldChange, topGene, pos=2, col="dodgerblue")
})
```
### Within the with function, only the baseMean and log2FoldChange values for the selected rows of res are used.

```{r}
plotMA(res, ylim = c(-5,5))
topGene <- rownames(res)[which.min(res$padj)]
with(res[topGene, ], {
  points(baseMean, log2FoldChange, col="dodgerblue", cex=2, lwd=2)
  text(baseMean, log2FoldChange, topGene, pos=2, col="dodgerblue")
})
```

```{r}
hist(res$pvalue[res$baseMean > 1], breaks = 0:20/20,
     col = "grey50", border = "white")
```

```{r}
library("genefilter")
topVarGenes <- head(order(rowVars(assay(vsd)), decreasing = TRUE), 20)
```

### Treatment status and cell line information are shown with colored bars at the top of the heatmap. Blocks of genes that covary across patients. Note that a set of genes in the heatmap are separating the N061011 cell line from the others, and there is another set of genes for which the dexamethasone treated samples have higher gene expression.

```{r}
mat  <- assay(vsd)[ topVarGenes, ]
mat  <- mat - rowMeans(mat)
anno <- as.data.frame(colData(vsd)[, c("cell","dex")])
pheatmap(mat, annotation_col = anno)
```
```{r}
qs <- c(0, quantile(resLFC1$baseMean[resLFC1$baseMean > 0], 0:6/6))
bins <- cut(resLFC1$baseMean, qs)
levels(bins) <- paste0("~", round(signif((qs[-1] + qs[-length(qs)])/2, 2)))
fractionSig <- tapply(resLFC1$pvalue, bins, function(p)
                          mean(p < .05, na.rm = TRUE))
barplot(fractionSig, xlab = "mean normalized count",
                     ylab = "fraction of small p values")

```

### In the following code chunk, we create bins using the quantile function, bin the genes by base mean using cut, rename the levels of the bins using the middle point, calculate the ratio of p values less than 0.05 for each bin, and finally plot these ratios (figure below).

```{r}
qs <- c(0, quantile(resLFC1$baseMean[resLFC1$baseMean > 0], 0:6/6))
bins <- cut(resLFC1$baseMean, qs)
levels(bins) <- paste0("~", round(signif((qs[-1] + qs[-length(qs)])/2, 2)))
fractionSig <- tapply(resLFC1$pvalue, bins, function(p)
                          mean(p < .05, na.rm = TRUE))
barplot(fractionSig, xlab = "mean normalized count",
                     ylab = "fraction of small p values")
```

```{r}
library("AnnotationDbi")
library("org.Hs.eg.db")
```

```{r}
columns(org.Hs.eg.db)
```

```{r}
ens.str <- substr(rownames(res), 1, 15)
res$symbol <- mapIds(org.Hs.eg.db,
                     keys=ens.str,
                     column="SYMBOL",
                     keytype="ENSEMBL",
                     multiVals="first")
res$entrez <- mapIds(org.Hs.eg.db,
                     keys=ens.str,
                     column="ENTREZID",
                     keytype="ENSEMBL",
                     multiVals="first")

```

### We can use the mapIds function to add individual columns to our results table. We provide the row names of our results table as a key, and specify that keytype=ENSEMBL. The column argument tells the mapIds function which information we want, and the multiVals argument tells the function what to do if there are multiple possible values for a single input value. 

```{r}
resOrdered <- res[order(res$pvalue),]
head(resOrdered)

```

### Exporting results

```{r}
resOrderedDF <- as.data.frame(resOrdered)[1:100, ]
write.csv(resOrderedDF, file = "results.csv")
```

### ReportingTools will automatically generate dynamic HTML documents, including links to external databases using gene identifiers and boxplots summarizing the normalized counts across groups.

```{r}
library("ReportingTools")
htmlRep <- HTMLReport(shortName="report", title="My report",
                      reportDirectory="./report")
publish(resOrderedDF, htmlRep)
url <- finish(htmlRep)
browseURL(url)
```

### file:///Users/jovan/Genomics-Course/report/report.html

```{r}
resGR <- lfcShrink(dds, coef="dex_trt_vs_untrt", type="apeglm", format="GRanges")
resGR
```

```{r}
ens.str <- substr(names(resGR), 1, 15)
resGR$symbol <- mapIds(org.Hs.eg.db, ens.str, "SYMBOL", "ENSEMBL")
```

```{r}
library("Gviz")
```

### The following code chunk specifies a window of 1 million base pairs upstream and downstream from the gene with the smallest p value. We create a subset of our full results, for genes within the window. We add the gene symbol as a name if the symbol exists and is not duplicated in our subset.

```{r}
window <- resGR[topGene] + 1e6
strand(window) <- "*"
resGRsub <- resGR[resGR %over% window]
naOrDup <- is.na(resGRsub$symbol) | duplicated(resGRsub$symbol)
resGRsub$group <- ifelse(naOrDup, names(resGRsub), resGRsub$symbol)
```

### We create a vector specifying if the genes in this subset had a low value of padj.

```{r}
status <- factor(ifelse(resGRsub$padj < 0.05 & !is.na(resGRsub$padj),
                        "sig", "notsig"))
```

### We create an axis track specifying our location in the genome, a track that will show the genes and their names, colored by significance, and a data track that will draw vertical bars showing the moderated log fold change produced by DESeq2, which we know are only large when the effect is well supported by the information in the counts.

```{r}
options(ucscChromosomeNames = FALSE)
g <- GenomeAxisTrack()
a <- AnnotationTrack(resGRsub, name = "gene ranges", feature = status)
d <- DataTrack(resGRsub, data = "log2FoldChange", baseline = 0,
               type = "h", name = "log2 fold change", strand = "+")
plotTracks(list(g, d, a), groupAnnotation = "group",
           notsig = "grey", sig = "hotpink")
```

```{r}
library("sva")
```

### Below we obtain a matrix of normalized counts for which the average count across samples is larger than 1

```{r}
dat  <- counts(dds, normalized = TRUE)
idx  <- rowMeans(dat) > 1
dat  <- dat[idx, ]
mod  <- model.matrix(~ dex, colData(dds))
mod0 <- model.matrix(~   1, colData(dds))
svseq <- svaseq(dat, mod, mod0, n.sv = 2)

```

```{r}
svseq$sv
```

```{r}
par(mfrow = c(2, 1), mar = c(3,5,3,1))
for (i in 1:2) {
  stripchart(svseq$sv[, i] ~ dds$cell, vertical = TRUE, main = paste0("SV", i))
  abline(h = 0)
 }
```

### we can see how well the SVA method did at recovering these variables

```{r}
ddssva <- dds
ddssva$SV1 <- svseq$sv[,1]
ddssva$SV2 <- svseq$sv[,2]
design(ddssva) <- ~ SV1 + SV2 + dex
```

### Finally, in order to use SVA to remove any effect on the counts from our surrogate variables, we simply add these two surrogate variables as columns to the DESeqDataSet and then add them to the design.

### Using RUV with DESeq2

```{r}
library("RUVSeq")
```

```{r}
set <- newSeqExpressionSet(counts(dds))
idx  <- rowSums(counts(set) > 5) >= 2
set  <- set[idx, ]
set <- betweenLaneNormalization(set, which="upper")
not.sig <- rownames(res)[which(res$pvalue > .1)]
empirical <- rownames(set)[ rownames(set) %in% not.sig ]
set <- RUVg(set, empirical, k=2)
pData(set)
```

```{r}
par(mfrow = c(2, 1), mar = c(3,5,3,1))
for (i in 1:2) {
  stripchart(pData(set)[, i] ~ dds$cell, vertical = TRUE, main = paste0("W", i))
  abline(h = 0)
 }
```

```{r}
ddsruv <- dds
ddsruv$W1 <- set$W_1
ddsruv$W2 <- set$W_2
design(ddsruv) <- ~ W1 + W2 + dex
```

### Time course experiments

```{r}
library("fission")
data("fission")
ddsTC <- DESeqDataSet(fission, ~ strain + minute + strain:minute)
```

### use a design formula that models the strain difference at time 0, the difference over time, and any strain-specific differences over time (the interaction term strain:minute).

```{r}
ddsTC <- DESeq(ddsTC, test="LRT", reduced = ~ strain + minute)
resTC <- results(ddsTC)
resTC$symbol <- mcols(ddsTC)$symbol
head(resTC[order(resTC$padj),], 4)
```


```{r}
fiss <- plotCounts(ddsTC, which.min(resTC$padj), 
                   intgroup = c("minute","strain"), returnData = TRUE)
fiss$minute <- as.numeric(as.character(fiss$minute))
ggplot(fiss,
  aes(x = minute, y = count, color = strain, group = strain)) + 
  geom_point() + stat_summary(fun.y=mean, geom="line") +
  scale_y_log10()
```

```{r}
resultsNames(ddsTC)
```

```{r}
res30 <- results(ddsTC, name="strainmut.minute30", test="Wald")
res30[which.min(resTC$padj),]
```

```{r}
betas <- coef(ddsTC)
colnames(betas)
```

```{r}
topGenes <- head(order(resTC$padj),20)
mat <- betas[topGenes, -c(1,2)]
thr <- 3 
mat[mat < -thr] <- -thr
mat[mat > thr] <- thr
pheatmap(mat, breaks=seq(from=-thr, to=thr, length=101),
         cluster_col=FALSE)
```
### The bottom set of genes show strong induction of expression for the baseline samples in minutes 15-60 (red boxes in the bottom left corner), but then have slight differences for the mutant strain (shown in the boxes in the bottom right corner).


### Session information, which reports the version numbers of R and all the packages used in this session
```{r}
sessionInfo()
```


