---
title: "Lab_9"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(stringr)
library(tidyverse)
```

```{r}
x <- c("apple", "banana", "pear")
str_view(x, "an")
```
```{r}
str_view(x, ".a.")
```
```{r}
# To create the regular expression, we need \\
dot <- "\\."

# But the expression itself only contains one:
writeLines(dot)
```

```{r}
# And this tells R to look for an explicit .
x <- c("abc", "a.c", "bef")
str_view(x, "a\\.c")
```
```{r}
x <- "a\\b"
writeLines(x)
```

```{r}
str_view(x, "\\\\")
```

### Exercise 14.3.1.1

### 1. "/" will result in the escape of the next adjacent character.

### "//" will result in essentially "/" as it will also result in an escape of the next character.

### "///" will result in the first two backslashes being expressed as literal backslashes in the expression but the third will escape the next character.



```{r}
x <- c("apple", "banana", "pear")
str_view(x, "^a")
```
```{r}
str_view(x, "a$")
```

```{r}
x <- c("apple pie", "apple", "apple cake")
str_view(x, "apple")
```
```{r}
str_view(x, "^apple$")

```
### 14.3.2.1 Exercises

###1. How would you match the literal string "$^$"?

```{r}
str_view(c("$^$", "ab$^$sfas"), "^\\$\\^\\$$", match = TRUE)
```

```{r}
# Look for a literal character that normally has special meaning in a regex
x <- c("abc", "a.c", "a*c", "a c")
str_view(x, "a[.]c")
```

```{r}
str_view(x, ".[*]c")
```

```{r}
str_view(x, "a[ ]")
```

```{r}
x <- c("grey", "gray")
str_view(x, "gr(e|a)y")
```
### 14.3.3.1 Exercises
### Create regular expressions to find all words that:
### Start with a vowel.

```{r}
str_subset(stringr::words, "^[aeiou]")
```


### That only contain consonants. (Hint: thinking about matching “not”-vowels.)

```{r}
str_subset(stringr::words, "[aeiou]", negate=TRUE)
```

###End with ed, but not with eed.

```{r}
str_subset(stringr::words, "[^e]ed$")
```


####End with ing or ise.
```{r}
str_subset(stringr::words, "i(ng|se)$")
```

```{r}
x <- "1888 is the longest year in Roman numerals: MDCCCLXXXVIII"
str_view(x, "CC?")
```

```{r}
str_view(x, "CC+")
```

```{r}
str_view(x, 'C[LX]+')
```

```{r}
str_view(x, 'C[LX]+')
```

```{r}
str_view(x, "C{2}")
```

```{r}
str_view(x, "C{2,}")
```

```{r}
str_view(x, "C{2,}")
```

```{r}
str_view(x, 'C{2,3}?')
```

```{r}
str_view(x, 'C[LX]+?')
```

### 14.3.4.1 Exercises

###Describe the equivalents of ?, +, * in {m,n} form.

### ?, {0,1} will match at the most 1

### +, {1,} will match one or more

### *, {0,} will match 0 or more

```{r}
str_view(fruit, "(..)\\1", match = TRUE)
```

### 14.3.5.1 Exercises
### Describe, in words, what these expressions will match:

### (.)\1\1 will match the same character that occurs three times in a row. This would be seen if "eee" was what you were trying to match.

### "(.)(.)\\2\\1" will match two characters and then the reverse of these two characters previous order. For example "xyyx".

### (..)\1 will match any two characters that are repeated. For example, "m8m8"

### "(.).\\1.\\1" will match a character and then right after any other character. This is then followed by the original character, another random character and then finally once again the original character. For example, "WxWyW"

### "(.)(.)(.).*\\3\\2\\1" will match three characters, a random amount of characters of any kind and then the original three characters but reversed. For example "jklsdfglkj"

### 14.4.1 Detect matches

```{r}
x <- c("apple", "banana", "pear")
str_detect(x, "e")

```


```{r}
# How many common words start with t?
sum(str_detect(words, "^t"))
```

```{r}
# What proportion of common words end with a vowel?
mean(str_detect(words, "[aeiou]$"))
```

```{r}
# Find all words containing at least one vowel, and negate
no_vowels_1 <- !str_detect(words, "[aeiou]")
```

```{r}
# Find all words consisting only of consonants (non-vowels)
no_vowels_2 <- str_detect(words, "^[^aeiou]+$")
identical(no_vowels_1, no_vowels_2)
```

```{r}
words[str_detect(words, "x$")]
```

```{r}
str_subset(words, "x$")

```

```{r}
df <- tibble(
  word = words, 
  i = seq_along(word)
)
df %>% 
  filter(str_detect(word, "x$"))
```

```{r}
x <- c("apple", "banana", "pear")
str_count(x, "a")
```

```{r}
# On average, how many vowels per word?
mean(str_count(words, "[aeiou]"))
```

```{r}
df %>% 
  mutate(
    vowels = str_count(word, "[aeiou]"),
    consonants = str_count(word, "[^aeiou]")
  )
```
```{r}
str_count("abababa", "aba")
```

```{r}
str_view_all("abababa", "aba")
```

### 14.4.1.1 Exercises

### For each of the following challenges, try solving it by using both a single regular expression, and a combination of multiple str_detect() calls.
### Find all words that start or end with x.
```{r}
words[str_detect(words, "^x|x$")]
```

```{r}
start_with_x <- str_detect(words, "^x")
end_with_x <- str_detect(words, "x$")
words[start_with_x | end_with_x]
```

### Find all words that start with a vowel and end with a consonant.
```{r}
str_subset(words, "^[aeiou].*[^aeiou]$") %>% head()
```

```{r}
start_with_vowel <- str_detect(words, "^[aeiou]")
end_with_consonant <- str_detect(words, "[^aeiou]$")
words[start_with_vowel & end_with_consonant] %>% head()
```

### Are there any words that contain at least one of each different vowel?

```{r}
pattern <-
  cross(rerun(5, c("a", "e", "i", "o", "u")),
    .filter = function(...) {
      x <- as.character(unlist(list(...)))
      length(x) != length(unique(x))
    }
  ) %>%
  map_chr(~str_c(unlist(.x), collapse = ".*")) %>%
  str_c(collapse = "|")
```

```{r}
words[str_detect(words, "a") &
  str_detect(words, "e") &
  str_detect(words, "i") &
  str_detect(words, "o") &
  str_detect(words, "u")]

```

### 14.4.2 Extract matches

```{r}
length(sentences)

```

```{r}
head(sentences)
```

```{r}
colours <- c("red", "orange", "yellow", "green", "blue", "purple")
colour_match <- str_c(colours, collapse = "|")
colour_match
```

```{r}
has_colour <- str_subset(sentences, colour_match)
matches <- str_extract(has_colour, colour_match)
head(matches)
```

```{r}
has_colour
```

```{r}
more <- sentences[str_count(sentences, colour_match) > 1]
str_view_all(more, colour_match)
```

```{r}
str_extract(more, colour_match)
```

```{r}
str_extract_all(more, colour_match)
```

```{r}
str_extract_all(more, colour_match, simplify = TRUE)

```

```{r}
x <- c("a", "a b", "a b c")
str_extract_all(x, "[a-z]", simplify = TRUE)
```

### 14.4.2.1 Exercises

### In the previous example, you might have noticed that the regular expression matched “flickered”, which is not a colour. Modify the regex to fix the problem.



```{r}
### This was the original color match pattern:


colours <- c("red", "orange", "yellow", "green", "blue", "purple")
colour_match <- str_c(colours, collapse = "|")
```

### Flickered was matched because it contains the word "red" inside of it. This can be fixed by putting boundaries at the end of each color to make sure that it does not match words like flickered. This is done by adding a \b.

```{r}
colour_match2 <- str_c("\\b(", str_c(colours, collapse = "|"), ")\\b")
colour_match2
#> [1] "\\b(red|orange|yellow|green|blue|purple)\\b"
more2 <- sentences[str_count(sentences, colour_match) > 1]

str_view_all(more2, colour_match2, match = TRUE)
```

### 14.4.3 Grouped matches

```{r}
noun <- "(a|the) ([^ ]+)"

has_noun <- sentences %>%
  str_subset(noun) %>%
  head(10)
has_noun %>% 
  str_extract(noun)

```

```{r}
has_noun %>% 
  str_match(noun)
```

```{r}
tibble(sentence = sentences) %>% 
  tidyr::extract(
    sentence, c("article", "noun"), "(a|the) ([^ ]+)", 
    remove = FALSE
  )
```
### 14.4.3.1 Exercises

### Find all words that come after a “number” like “one”, “two”, “three” etc. Pull out both the number and the word.

```{r}
numword <- "\\b(one|two|three|four|five|six|seven|eight|nine|ten) +(\\w+)"
sentences[str_detect(sentences, numword)] %>%
  str_extract(numword)
```

### 14.4.4 Replacing matches

```{r}
x <- c("apple", "pear", "banana")
str_replace(x, "[aeiou]", "-")
```

```{r}
str_replace_all(x, "[aeiou]", "-")
```

```{r}
x <- c("1 house", "2 cars", "3 people")
str_replace_all(x, c("1" = "one", "2" = "two", "3" = "three"))
```

```{r}
sentences %>% 
  str_replace("([^ ]+) ([^ ]+) ([^ ]+)", "\\1 \\3 \\2") %>% 
  head(5)
```

### 14.4.4.1 Exercises

### Replace all forward slashes in a string with backslashes.

```{r}
str_replace_all("past/present/future", "/", "\\\\")
#> [1] "past\\present\\future"
```

### 14.4.5 Splitting

```{r}
sentences %>%
  head(5) %>% 
  str_split(" ")
```

```{r}
"a|b|c|d" %>% 
  str_split("\\|") %>% 
  .[[1]]
```

```{r}
sentences %>%
  head(5) %>% 
  str_split(" ", simplify = TRUE)
```

```{r}
fields <- c("Name: Hadley", "Country: NZ", "Age: 35")
fields %>% str_split(": ", n = 2, simplify = TRUE)
```

```{r}
x <- "This is a sentence.  This is another sentence."
str_view_all(x, boundary("word"))
```

```{r}
str_split(x, " ")[[1]]
```

```{r}
str_split(x, boundary("word"))[[1]]
```


### 14.4.5.1 Exercises

### Split up a string like “apples, pears, and bananas” into individual components.

```{r}
x <- c("apples, pears, and bananas")
str_split(x, ", +(and +)?")[[1]]
#> [1] "apples"  "pears"   "bananas"

```


